---
title: "Growth curve analysis with mixed modeling (use latent factors)"
author: "Yen Low"
date: "06/19/2015"
output: 
  html_document:
    self_contained: no
---
  
### Setup
Set paths and load dependencies
```{r,echo=FALSE}
RScriptPath="/mnt/hgfs/scripts"
ProjectPath="/mnt/hgfs/projects"
setwd(paste(ProjectPath,"/insight/models/hlm_fa",sep=""))
getwd()
 
source(paste(RScriptPath,"/R/utils.R",sep=""))
source(paste(RScriptPath,"/R/forestplot.R",sep=""))

require(nlme)
require(lattice)
require(latticeExtra)
```

### Load data
Use data matrix generated from FA and processed in ../glmmlasso_fa/datamat.Rmd
```{r,echo=FALSE}
load(file="../glmmlasso_fa/datamat.RData",verbose=T)
summary(tstMat$totalRevenue)
```
---


### Linear mixed modeling
*Level 1:* monthly revenue  
*Level 2:* store level  
*Significant factors:* timeOrder, MedicalRxStore, Uninsured, Commute, OtherBiz, Pollution, ProfServices
#### Intercept as outcome model
```{r}
#Formulate model
varString=paste(colnames(trnMat)[2:16],collapse=" + ")
formulaString=paste0("totalRevenue ~ timeOrder + ",varString)
formulaString
mod_intAsOutcome=lme(as.formula(formulaString),
                    random=~timeOrder|storeID,data=trnMat,
                    na.action=na.omit,control=list(opt="optim"))
summary(mod_intAsOutcome)
```


#### Slope as outcome model
*Significant factors:* timeOrder, MedicalRxStore, Uninsured, Commute, OtherBiz, Pollution, ProfServices
timeOrder:HealthCost, timeOrder:ProfServices, timeOrder:Alcholism
```{r}
varString2=paste0("timeOrder*",colnames(trnMat)[2:16],collapse=" + ")
formulaString2=paste0("totalRevenue ~ ",varString2)
formulaString2

mod_slopeAsOutcome=lme(as.formula(formulaString2),
                        random=~timeOrder|storeID,data=trnMat,
                        na.action=na.omit,control=list(opt="optim"))
summary(mod_slopeAsOutcome)
```


#### Best slope as outcome model
*Significant factors:* timeOrder, MedicalRxStore, Uninsured, Commute, Pollution, OtherBiz, ProfServices, timeOrder:HealthCost, timeOrder:ProfServices, timeOrder:Alcholism
```{r}
formulaString3=paste0(formulaString," + timeOrder*HealthCost + timeOrder*ProfServices + timeOrder*Alcholism")
mod_bestslopeAsOutcome=lme(as.formula(formulaString3),
                        random=~timeOrder|storeID,data=trnMat,
                        na.action=na.omit,control=list(opt="optim"))
summary(mod_bestslopeAsOutcome)
```


#### Update with polynomial time
*Significant factors:* timeOrder, timeOrder^2, timeOrder^3, MedicalRxStore, Uninsured, Commute, Pollution, ProfServices, Other Biz, timeOrder:HealthCost, timeOrder:ProfServices, timeOrder:Alcholism
```{r}
formulaString4=paste0("totalRevenue ~ timeOrder + I(timeOrder^2) + I(timeOrder^3) + ",varString," + timeOrder*HealthCost + timeOrder*ProfServices + timeOrder*Alcholism")
formulaString4
mod_polytime=lme(as.formula(formulaString4),
                        random=~timeOrder|storeID,data=trnMat,
                        na.action=na.omit,control=list(opt="optim"))
summary(mod_polytime)
```


#### Update with ARIMA
*Significant factors:* timeOrder, timeOrder^2, timeOrder^3, MedicalRxStore, Uninsured, Commute, Pollution, ProfServices, Other Biz, timeOrder:HealthCost, timeOrder:ProfServices
```{r}
mod_cortime=lme(as.formula(formulaString4),
                        random=~timeOrder|storeID,data=trnMat,
                        correlation=corExp(),
                        na.action=na.omit,control=list(opt="optim"))
summary(mod_cortime)
attributes(summary(mod_cortime))

d=as.data.frame(summary(mod_cortime)$tTable[,1:2])
colnames(d)[1]="median"
d$lowlim=d$median-1.96*d$Std.Error
d$upplim=d$median+1.96*d$Std.Error

write.table(d,file="beta_HLM.txt",sep="\t",quote=F,col.names=NA)
save(mod_cortime,d,file="hlm_fa.RData")
#load(file="hlm_fa.RData",verbose=T)

forestplot(d[nrow(d):2,],orderoflevels=NULL,xlab="Beta coefficients",ylab="",xmin=-2000, xmax=2000, xgap=500, sortByLabels=FALSE)
```



### Compare models with LRT
Best slope as outcome model is the best model
```{r}
anova(mod_slopeAsOutcome,mod_intAsOutcome)
anova(mod_bestslopeAsOutcome,mod_slopeAsOutcome)
anova(mod_bestslopeAsOutcome,mod_intAsOutcome)
anova(mod_polytime,mod_bestslopeAsOutcome)
anova(mod_cortime, mod_polytime)
```


### Validate best HLM model
#### On training data
```{r trnTS, echo=FALSE}
attributes(mod_cortime)
trnMatpred=trnMat
trnMatpred$predictedRev=fitted(mod_cortime)
selStore=sample(unique(trnMat$storeID),20)
actualRev=xyplot(totalRevenue~timeOrder|storeID,data=trnMat[trnMat$storeID %in% selStore,],col="red",type="p")
predRev=xyplot(predictedRev~timeOrder|storeID,data=trnMatpred[trnMatpred$storeID %in% selStore,],col="black",type="l")
actualRev + as.layer(predRev)
```
tmp=dcast(trnMatpred, timeOrder ~ storeID, value.var="totalRevenue")
plot(tmp[,1:2],type="l",ylim=c(0,50000),xlab="Month",ylab="Revenue ($)")
for (i in sample(1:ncol(tmp),10)) points(tmp[,c(1,i)],type="l",col=i)

#### On test data
```{r tstTS, echo=FALSE}
tstMat$storeID=as.character(tstMat$storeID)
meanRev=tapply(tstMat$totalRev,as.character(tstMat$storeID),mean,na.rm=T)
predictedRev=tapply(predict(mod_cortime, newdata=tstMat, level=0),tstMat$storeID,mean,na.rm=T)


resid=tstMatpred$predictedRev-tstMatpred$totalRevenue
mse=mean(resid^2,na.rm=T)
mse
1-mse/var(tstMatpred$totalRevenue)
plot(meanRev,predictedRev)
abline
cor(meanRev,predictedRev,method="spearman")^2
qqplot(meanRev,predictedRev)
tmp=coefficients(mod_cortime)


save(mod_cortime,d,tstMatpred,file="hlm_fa.RData")
```


## TODO
1. consider 3rd level (county)
