---
title: "Growth curve analysis with regularized mixed modeling (using latent factors)"
author: "Yen Low"
date: "05-Jul-15"
output: 
  html_document:
    self_contained: no
---
  
## Setup
Set paths and load dependencies
```{r,echo=FALSE}
RScriptPath="/mnt/hgfs/scripts"
ProjectPath="/mnt/hgfs/projects"
setwd(paste(ProjectPath,"/insight/models/glmmlasso_fa",sep=""))
getwd()

source(paste(RScriptPath,"/R/utils.R",sep=""))

require(glmmLasso)
require(lattice)
```

## Load data
Get monthly revenue (18-mth time series for 1487 stores)
```{r,echo=FALSE,message=FALSE}
load(file="../glmmlasso/datamat.RData",verbose=T)
totRev=datamat[,c("totalRevenue","timeOrder","storeID")]
rm(trnMat,tstMat,datamat,columnMean,columnSD)
```

Load store data (1109 stores, 15 latent variables)
```{r}
load(file="../fa/FA.RData",verbose=T)
dim(factorScores)
head(factorScores,3)
```

## Make feature matrix
Set time order for nlme
```{r,echo=FALSE}
factorScores=as.data.frame(factorScores)
factorScores$storeID=as.factor(rownames(factorScores))
datamat=merge(factorScores,totRev,by="storeID",sort=T)
datamat=datamat[order(datamat$storeID,datamat$timeOrder),]
dim(datamat)  #19792    18
summary(datamat$totalRevenue)
```

## Partition data into training and test sets
Get test store IDs (extract 133 storeIDs from 10% test set used to validate random forest model).  
Remove 134 stores with short time series (13 < mths)  
999 stores, 17840 rows for training  
125 stores, 2074 rows for testing (10%)  
```{r, echo=T}
#Get test store IDs (extract from test set used to validate random forest model)
trnMat=datamat[!(datamat$storeID %in% tstStore),] #17808 x 19
tstMat=datamat[datamat$storeID %in% tstStore,]  #1981 x 19

length(unique(trnMat$storeID))  #998 stores, 17824 rows for training
length(unique(tstMat$storeID))  #111 stores, 1968 rows for testing (10%)

summary(trnMat$totalRevenue)
summary(tstMat$totalRevenue)

save(trnMat,tstMat,tstStore,datamat,file="datamat.RData")
```

No need to rescale factor scores  
No missing values in factor scores



