---
title: "Monthly Sales Growth"
author: "Yen Low"
date: "06/11/2015"
output: 
  html_document:
    self_contained: no
---
  
### Setup
Set paths and load dependencies
```{r,echo=FALSE}
RScriptPath="/mnt/hgfs/scripts"
ProjectPath="/mnt/hgfs/projects"
setwd(paste(ProjectPath,"/insight/models/timeseries",sep=""))
getwd()

source(paste(RScriptPath,"/R/utils.R",sep=""))
source(paste(RScriptPath,"/R/localMySQL.R",sep=""))
source(paste(RScriptPath,"/R/hampel.R",sep=""))
source(paste(RScriptPath,"/R/growth.R",sep=""))

require(RMySQL)
require(astsa)
require(TTR)
#require(plyr)
```

### Load data
Get monthly sales
```{r,echo=FALSE,message=FALSE}
con = dbConnect(drv, username=username, password=passwd)
totRev=dbGetQuery(con,"SELECT * FROM testdb.totsales_month")
dbDisconnect(con)
```

Check number of stores (1487)
```{r,echo=TRUE,message=TRUE}
#1487 stores
storeIDs=unique(totRev$customer_identifier)
length(storeIDs)
```

### Data Maniputation
Convert to dataframe and time series object
1487 stores over 19 mths
```{r,echo=FALSE,message=TRUE}
tsDf=xtabs(totalRevenue ~ customer_identifier + yr_month, totRev)
head(tsDf,3)

#Then put into a time series object (as ts object from astsa package)
tsfunc<-function(x){
  yr_mth=names(x)
  yr=substr(min(yr_mth),1,4)
  mth=substr(min(yr_mth),5,6)
  ans=ts(x,start=as.integer(c(yr,mth)),freq=12)
  return(list(ans))
}
tsList=apply(tsDf,1,tsfunc)
#x=tsList[[1]]
```

View time series
```{r,echo=FALSE,message=TRUE}
plot(tsList[[1]][[1]])
```


### Calculate growth rates
Growth rate = (x(tn) - x(tn-1))x(tn-1)
Average growth rates are calculated from trimmed means of at least 3 values
Outliers are trimmed by Hampel's test (>3 SD from median)
```{r,echo=FALSE,message=TRUE}
#Monthly growth
growth_mth=lapply(tsList, function(x) growthrate(x[[1]], n=1))
#x=growth_mth[[1]]
aveMthGrowth=sapply(growth_mth,trimmedMean_hampel,RemoveNAs=T,na.rm=T,minlen=3)

#Quarterly growth
growth_qtr=lapply(tsList, function(x){
  newts=to.quarterly(as.xts(x[[1]]),indexAt="yearmon")[,4]
  ans=growthrate(newts[-nrow(newts),], n=1)
  return(ans)
})
aveQtrGrowth=sapply(growth_qtr,trimmedMean_hampel,RemoveNAs=T,na.rm=T,minlen=3)
# x=growth_qtr[[1]]
# temp=list()
# for(i in 1:length(growth_qtr)){
#   x=growth_qtr[[i]]
#   temp[[i]]=trimmedMean_hampel(x)
# }

#Annual growth
growth_yr=lapply(tsList, function(x){
  newts=to.yearly(as.xts(x[[1]]),indexAt="yearmon")[,4]
  ans=growthrate(newts[-nrow(newts),], n=1)
  return(ans)
})
aveYrGrowth=sapply(growth_yr,trimmedMean_hampel,RemoveNAs=T,na.rm=T,minlen=1)
```

View histograms of growth rates
#### Monthly growth
View monthly growth of first store
```{r, echo=FALSE}
growth_mth[[1]]
#sort(aveMthGrowth)
lim=0.5
tmp=summary(aveMthGrowth)
tmp
hist(aveMthGrowth[abs(aveMthGrowth)<lim],breaks=seq(-lim,lim,by=0.1))
abline(v=tmp["Median"],lty=2)
abline(v=tmp["Mean"])
```


#### Quarterly growth
View quarterly growth of first store
```{r,echo=FALSE}
growth_qtr[[1]]
#sort(aveQtrGrowth)
lim=1
tmp=summary(aveQtrGrowth[abs(aveQtrGrowth)<lim],)
tmp
hist(aveQtrGrowth[abs(aveQtrGrowth)<lim],breaks=seq(-lim,lim,by=0.1))
abline(v=tmp["Median"],lty=2)
abline(v=tmp["Mean"])
```


#### Quarterly growth
View quarterly growth of first store
```{r,echo=FALSE}
growth_yr[[1]]
#sort(aveYrGrowth)
tmp=summary(aveYrGrowth[aveYrGrowth>-2 & aveYrGrowth<4],)
tmp
hist(aveYrGrowth[aveYrGrowth>-2 & aveYrGrowth<4],breaks=seq(-2,4,by=0.1))
abline(v=tmp["Median"],lty=2)
abline(v=tmp["Mean"])
legend("topright", c("Mean","Median"),lty=c(1,2))
```

### Check for problematic stores
Some problemantic stores: 983030 182941, 979649, 086800, 015305
(not enough values, extreme growth rates, etc)
```{r,echo=TRUE}
tsList[["015305"]]
growth_yr[["015305"]]
aveYrGrowth["015305"]
```

### Set thresholds for high growth
Min 5% average monthly growth, 10% quarterly growth, 25% yearly growth
```{r,echo=TRUE}
sum(na.omit(aveMthGrowth>0.05 & is.finite(aveMthGrowth)))   #368
sum(na.omit(aveQtrGrowth>0.1 & is.finite(aveQtrGrowth)))  #417
sum(na.omit(aveYrGrowth>0.25 & is.finite(aveYrGrowth)))     #499
```

### Save growth rates and average growth rates into data frames
Dataframe of average growth rate
```{r,echo=TRUE}
aveGrowthDf=as.data.frame(cbind(aveMthGrowth,aveQtrGrowth,aveYrGrowth))
head(aveGrowthDf,3)
```

Indicator dataframe of high average growth rate using the above thresholds
Min 5% average monthly growth, 10% quarterly growth, 25% yearly growth
```{r,echo=FALSE}
highGrowthDf=as.data.frame(matrix(0,nrow=nrow(aveGrowthDf),ncol=ncol(aveGrowthDf)))
rownames(highGrowthDf)=rownames(aveGrowthDf)
colnames(highGrowthDf)=c("highMthGrowth","highQtrGrowth","highYrGrowth")
highGrowthDf[which(aveGrowthDf$aveMthGrowth>0.05),"highMthGrowth"]=1
highGrowthDf[which(aveGrowthDf$aveQtrGrowth>0.1),"highQtrGrowth"]=1
highGrowthDf[which(aveGrowthDf$aveYrGrowth>0.25),"highYrGrowth"]=1
highGrowthDf[!is.finite(as.matrix(aveGrowthDf))]=NA
head(highGrowthDf,3)
```

Check for NA values
```{r,echo=TRUE}
sum(is.na(highGrowthDf))
```

Check pairwise concordance among average monthly, quarterly and yearly growth rates
```{r,echo=FALSE}
pairIDs=combn(1:3,2)
for(i in 1:ncol(pairIDs)) print(table(highGrowthDf[,pairIDs[,i]]))
```

View store with highest average monthly growth
```{r,echo=TRUE}
bestStore=names(which.max(aveMthGrowth))
aveMthGrowth[bestStore]
growth_mth[[bestStore]]
tsList[[bestStore]]
#png("bestQtrGrowth.png")
plot(growth_mth[[bestStore]], main="Store with highest average monthly growth", ylab="monthly growth ()")
#dev.off()
```

Save time series objects and data frames into data/revGrowth.RData
```{r}
save(tsList,highGrowthDf,aveGrowthDf,
     growth_mth,growth_qtr,growth_yr,file=paste0(ProjectPath,"/insight/data/revGrowth.RData"))
```

