---
title: "Store Characteristics"
author: "Yen Low"
date: "06/11/2015"
output: 
  html_document:
    self_contained: no
---

### Setup
Set paths and load dependencies
```{r setOptions,echo=FALSE, message=FALSE}
#set paths
RScriptPath="/mnt/hgfs/scripts"
ProjectPath="/mnt/hgfs/projects"
setwd(paste(ProjectPath,"/insight/models/descriptive",sep=""))
getwd()

#load dependencies
source(paste(RScriptPath,"/R/utils.R",sep=""))
source(paste(RScriptPath,"/R/localMySQL.R",sep=""))

require(RMySQL)
suppressPackageStartupMessages(require(googleVis))
require(choroplethrZip)  #provides zip.regions data for mapping zip to county
op=options(gvis.plot.tag="chart")

#get zip to state mappings
data(zip.regions)
zip.regions$county.fips=formatC(zip.regions$county.fips.numeric,width=5,format='f',digits=0,flag='0')
```

### Load data
Load store charactistics (type, zip) from MySQL DB.
```{r,echo=FALSE}
#get store type
con = dbConnect(drv, username=username, password=passwd,"testdb")
storeDf=dbGetQuery(con,"SELECT * FROM stores")
dbDisconnect(con)
```

```{r,echo=TRUE}
head(storeDf,3)
```

### Number of stores by type
```{r,echo=FALSE}
numStoresByType=sort(table(storeDf$busType,useNA="ifany"))
rev(numStoresByType)

par(mar=c(2,15,2,1))
bp=barplot(numStoresByType,horiz=T,log="x",las=1,cex.names=0.8,
           main="Number of stores by category")
text(1,bp,numStoresByType,cex=0.8,pos=4)
```

#### Collapse store types into fewer, more meaningful groups
```{r}
storeDf$bizType=NA
storeDf$bizType[grep("drug|retail",storeDf$busType,ignore.case=T)]="retail"
storeDf$bizType[grep("hospital|center|physician",storeDf$busType,ignore.case=T)]="medical"
storeDf$bizType[grep("wholesale|internal|closed pharmacy|customer|home",storeDf$busType,ignore.case=T)]="others"
newNumStoresByType=sort(table(storeDf$bizType,useNA="ifany"))

par(mar=c(2,5,2,1))
bp=barplot(newNumStoresByType,horiz=T,log="x",las=1,cex.names=1,
           main="Number of stores by category")
text(100,bp,newNumStoresByType,cex=1,pos=4)

```

Make indicator variables out of collapsed business types
```{r}
#tmp=model.matrix(~ bizType, data=storeDf[1:100,])
storeTypeAsFeature=as.data.frame.matrix(xtabs(~ storeID + bizType, data=storeDf))
head(storeTypeAsFeature,3)
storeDf=merge(storeDf,storeTypeAsFeature,by.x="storeID",by.y=0)
head(storeDf,3)
```


### Number of stores by region
```{r,echo=FALSE}
#count stores by state
store_state=as.data.frame(table(storeDf$state))
colnames(store_state)=c("state","Stores")

#count stores by zip
store_zip=as.data.frame(table(storeDf$zip))
colnames(store_zip)=c("zip","Stores")

#map zip to county
store_county=store_zip
#2347 stores
tmp=merge(store_county,zip.regions[,c("region","county.name","county.fips")], by.x="zip", by.y="region")
#don't double count same store in 2 counties (cos 1 zip can map to 2 counties)
#delete the second county
tmp=tmp[order(tmp$zip,tmp$county.fips),]
tmp=tmp[!duplicated(tmp$zip),] #1427 stores
store_county=aggregate(Stores~county.fips,tmp,sum)   #752 counties
colnames(store_county)=c("county","Stores")
```

Preview stores by region
```{r}
head(store_state,3)
head(store_county,3)
head(store_zip,3)
```

#### State
```{r results='asis'}
m=gvisGeoChart(store_state,
               locationvar="state", colorvar="Stores",
               options=list(region="US", displayMode="regions", 
                            resolution="provinces",
                            width=1000, height=600,
                            colorAxis="{colors:['#FFFFFF', '#0000FF']}"
               )) 
plot(m)
```

#### Output data files
saves dataframes to  data/storeChar.RData
also output number of stores by county to "store_county.txt"
```{r}
write.table(store_county,file="store_county.txt",sep="\t",row.names=F,quote=F)

save(store_state,store_county,store_zip,storeDf,file="../../data/storeChar.RData")
```
