---
title: "Predict which store is fast growing using a Random Forest classifier"
author: "Yen Low"
date: "06/16/2015"
output: 
  html_document:
    self_contained: no
---
  
### Setup
Set paths and load dependencies
```{r,echo=FALSE}
RScriptPath="/mnt/hgfs/scripts"
ProjectPath="/mnt/hgfs/projects"
setwd(paste(ProjectPath,"/insight/models/rf_fastVsSlow",sep=""))
getwd()

source(paste(RScriptPath,"/R/utils.R",sep=""))

require(randomForest)
require(choroplethrZip) 
```

### Load data
Get indicator variables identifying stores as fast or slow growing (1487 stores)
```{r,echo=FALSE,message=FALSE}
load(file=paste0(ProjectPath,"/insight/data/revGrowth.RData"),verbose=T)  #response: sales growthload
head(highGrowthDf,3)
```

Load store features (1437 stores)
```{r}
load(file=paste0(ProjectPath,"/insight/data/dsStore.RData"),verbose=T)  #store level
str(dsStore)
```

Load data matrix and the test set for random forest
```{r, echo=FALSE}
dsMat=as.data.frame(merge(highGrowthDf,dsStore,by.x=0,by.y="storeID",sort=T))
str(dsMat)
save(dsMat,file="dsMat.RData")
#load(file="dsMat.RData",verbose=T)
```

### Test set for model validation
Set aside 10% test set for validation
Save this test set for validating other models
```{r}
testID=sample(nrow(dsMat),0.1*nrow(dsMat))
#ensure no NA or missing values (do zero imputation)
sum(is.na(dsMat))
dsMat[is.na(dsMat)]=0
sum(is.na(dsMat))
testStores=dsMat$Row.names[testID]
save(dsMat,testID,file="dsMat.RData")
save(testStores,file="testStores.RData")
#load(file="dsMat.RData",verbose=T)
```

### Random forest
Acceptable prediction performance (OOB=0.27)
Consider adding more features or model average monthly growth instead of fast vs slow monthly sales
```{r}
#drop variables that you don't wish to model
excludeVarID=grep("Row.names|high|zip|city|fip|county|^state$|Type$",colnames(dsMat))
colnames(dsMat)[excludeVarID]
#ensure that response highMthGrowth is a factor to run randomForest in classification mode
dsMat$highMthGrowth=as.factor(dsMat$highMthGrowth)
mod_rf=randomForest(x=dsMat[-testID,-excludeVarID],y=dsMat$highMthGrowth[-testID],
                    xtest=dsMat[testID,-excludeVarID],ytest=dsMat$highMthGrowth[testID])
mod_rf
```

#### Variable importance in random forest model
```{r varImpPlot, echo=FALSE, out.width=1000, out.height=9000}
imptVar=as.matrix(importance(mod_rf))
imptVar=sort(imptVar[,"MeanDecreaseGini"],decreasing=T)
par(mar=c(2,20,2,0))
barplot(imptVar[50:1],horiz=T,las=1,log="x",cex.names=0.6, main="Variable importance")
```

Less important variables (in decreasing importance)
```{r echo=FALSE}
as.matrix(imptVar[51:length(imptVar)])
```

Save RF model
```{r}
write.table(as.matrix(imptVar),file="imptVar.txt",col.names=F,quote=F,sep="\t")
save(mod_rf,imptVar,file="mod.RData")
```

## TODO
1. consider other features (drugs)
2. consider time-varying features (e.g. emp rate)
3. consider continuous random forest (modeling average monthly growth)
4. consider polynomials(t)
5. consider lasso logistic regression
