---
title: "Predict revenue at any timepoint using Random Forest regression (assumes revenue is independent within store's time serie)"
author: "Yen Low"
date: "06/12/2015"
output: 
  html_document:
    self_contained: no
---
  
### Setup
Set paths and load dependencies
```{r,echo=FALSE}
RScriptPath="/mnt/hgfs/scripts"
ProjectPath="/mnt/hgfs/projects"
setwd(paste(ProjectPath,"/insight/models/rf_timeseries",sep=""))
getwd()

source(paste(RScriptPath,"/R/utils.R",sep=""))
source(paste(RScriptPath,"/R/hampel.R",sep=""))
#source(paste(RScriptPath,"/R/growth.R",sep=""))

require(randomForest)
```

### Load data
Get average quarterly growth rates of stores (1487 stores)
```{r,echo=FALSE,message=FALSE}
load(file="../glmmlasso_fa/datamat.RData",verbose=T)


head(aveGrowthDf,3)
#remove stores with NA ave growth
aveGrowthDf=aveGrowthDf[!is.na(aveGrowthDf$aveQtrGrowth),]
#check 1462 remaining stores do not have NA average monthly growth
sum(is.na(aveGrowthDf$aveQtrGrowth))
```

#### Remove outliers
Histogram indicate outliers. Remove using Hampel's rule
```{r}
hist(aveGrowthDf$aveQtrGrowth,100)
#out=boxcox(lm((aveGrowthDf$aveQtrGrowth +10)~1))
#range(out$x[out$y > max(out$y)-qchisq(0.95,1)/2])
trimmedVal=trimByHampel(log10(aveGrowthDf$aveQtrGrowth+2),t=3)
#trimmed 188 outliers
length(aveGrowthDf$aveQtrGrowth)-length(trimmedVal)
hist(trimmedVal,100)
```

Load store features (1602 stores)
```{r}
load(file=paste0(ProjectPath,"/insight/data/dsStore.RData"),verbose=T)  #store level
head(dsStore,3)
```

Load data matrix and the test set for random forest
```{r, echo=FALSE}
load(file="dsMat.RData",verbose=T)
dsMat=as.data.frame(merge(aveGrowthDf,dsStore,by.x=0,by.y="storeID",sort=T))
dsMat$empRate=dsMat$nEmployed/dsMat$population

#zero-impute NA values
dsMat[is.na(dsMat)]=0
#check that no NA values remain
sum(is.na(dsMat))
head(dsMat,3)
save(dsMat,file="dsMat.RData")
```

### Test set for model validation
Set aside 10% test set for validation
Save this test set for validating other models
```{r}
testID=sample(nrow(dsMat),0.1*nrow(dsMat))
save(dsMat,testID,file="dsMat.RData")
```


### Random forest
Poor prediction performance (OOB=0.27)
Consider adding more features or model average monthly growth instead of fast vs slow monthly sales
```{r}
#drop variables that you don't wish to model
excludeVarID=grep("Row.names|^ave.+Growth$|zip|busType|nEmployed",colnames(dsMat))
#ensure that response aveMthGrowth is numeric to run randomForest in regression mode
dsMat$aveQtrGrowth=as.numeric(dsMat$aveQtrGrowth)
mod_rf=randomForest(x=dsMat[-testID,-excludeVarID],y=log10(2+dsMat$aveQtrGrowth[-testID]),
                    xtest=dsMat[testID,-excludeVarID],ytest=log10(2+dsMat$aveQtrGrowth[testID]),
                    importance=T)
mod_rf
#get model R^2
mod_rf$test$rsq[mod_rf$ntree]
#get model MSE
mod_rf$test$mse[mod_rf$ntree]
qqplot(log10(2+dsMat$aveQtrGrowth[testID]),mod_rf$test$predicted)

```

#### Variable importance in random forest model
```{r varImpPlot, echo=FALSE, out.width=1000, out.height=3000}
imptVar=as.matrix(importance(mod_rf))
imptVar=imptVar[,1]
par(mar=c(1,20,2,0))
barplot(sort(imptVar),horiz=T,las=1,cex.names=0.7, main="Variable importance")
```


Save RF model
```{r}
save(mod_rf,imptVar,file="mod.RData")
```

## TODO
1. consider other features (drugs, health indicators)
2. re-engineer features (e.g. purchasing power rather than income)
4. consider polynomials(t)
5. consider lasso logistic regression
